# Data exploration on raw data for Amazon project

### Setting 
```{r}
setwd("/Users/aszostek/Projects/Kaggle/Amazon")
source("./Utils/utils.R")
train <- read.csv(file="./Data/train.csv")
test <- read.csv(file="./Data/test.csv")

```

## Metric in this competition is AUC interpreted as:
* You do your classification and classify as 0 or 1
* Repeat - randomly choose one observation from set classified as 0 and one from set classified as 1
* Check if classification was true for both observations
* The area under the curve is the percentage of pairs with both observations classified correctly.
* That is why if you made a random classification, this will be equal ~0.5




## How many resources there are?
```{r}
length(unique(train[[2]]))
```
7k resources!!! that is a lot! In particular that there are only .. rows

```{r}
nrow(train)
```
How many managers?
```{r}
length(unique(train[[3]]))
length(unique(train[[4]]))
length(unique(train[[5]]))
length(unique(train[[6]]))
length(unique(train[[7]]))
length(unique(train[[8]]))
length(unique(train[[9]]))
length(unique(train[[10]]))
```
What are the most common resources requested?
```{r}
tb <- table(train[[2]])
cou <- factor(train[[2]], levels = names(tb[order(tb, decreasing = F)]))
#table(cou)
```
 Were they accepted or not? What are the resources that are rejected the most? Lets look for example resource 4675
```{r}
res = train[train$RESOURCE == 4675,]
table(res[[1]])
```
It seems that in this resource, majority requests are accepted!
```{r}
res = train[train$RESOURCE == 79092,]
table(res[[1]])
```
Here too
```{r}
res = train[train$RESOURCE == 75078,]
table(res[[1]])
```
and here too
```{r}
res = train[train$RESOURCE == 25993,]
table(res[[1]])
```
```{r}
res = train[train$RESOURCE == 3853,]
table(res[[1]])
```
### Question:
Are there any resources in test file that are not in training set? 
```{r}
a1 <- unique(train[[2]])
a2 <- unique(test[[2]])
length(setdiff(a1,a2))
```
#### Answer: yes there are more than 2000 of them!

### Question: What would be the accuracy if I set all actions to one for the training set?
```{r}
table(train[[1]])
a1 <- length(train[train$ACTION == 1,])
1-a1/nrow(train)
```
#### Answer: It seems like a pretty good accuracy to me :)

### Question: Is the data normalized?
```{r}
par(mfrow=c(1,2))
plot_summary(train,main="Training set")
plot_summary(test,main="Test set")
```
#### Answer:
* It seems it is
* For variables 4-10 the means are similar
* Variables 7 and 10 have very high minima compared to variables 4, 6, 8, 9
* the 1st quantile of variables 4-10 is almost identical
* also the median is almost identical
* I suspect that there was some effort to assign the numbers to categories in a sensible way!


### Question: Which managers accept resources more than others
* red - action 1
* black - action 0
```{r}
par(xpd=FALSE)
plot(train[train$ACTION==1,c(2,3)],col="red",pch=19,ylab="MANAGER")
points(train[train$ACTION==0,c(2,3)],col="black",pch=19)
abline(h=140000)
abline(h=50000)
abline(h=70000)
abline(h=5000)
```
#### Answer:
* Managers with number > 140,000 seem to accept all/majority requests!

### Question: What is the ratio of accepted vs rejected for given resource? or manager? Lets find the most problematic resources and managers!

```{r}
par(mfrow = c(1,2))
res <- unique(train[[2]])
v = c()
for (i in res)
{
  
  acc1 <- length(train[train$ACTION==1 & train$RESOURCE == i,1])
  acc0 <- length(train[train$ACTION==0 & train$RESOURCE == i,1])
  v <- append(v,as.numeric(acc0/(acc1+1)))
  #print(acc0/(acc1+1))
}
plot(v ~ res,xlim=c(0,150000),pch=19,xlab="Resource number",ylab="Nb of acceptances / Nb of rejections")

res <- unique(train[[3]])
v = c()
for (i in res)
{
  #acc1 <- length(train[train$ACTION==1 & train$MGR_ID == i,1])
  acc0 <- length(train[train$ACTION==0 & train$MGR_ID == i,1])
  acc <- length(train[train$MGR_ID == i,1])
  v <- append(v,as.numeric(acc0/acc))
  #print(acc0/(acc1+1))
}
plot(v ~ res,xlim=c(0,150000),pch=19,xlab="Manager number",ylab="Ratio of rejections to all requests")

```
### Question: How does Role title correlate with acceptance?
```{r}
par(xpd=FALSE)
plot(train[train$ACTION==1,c(2,7)],col="red",pch=19,ylab="ROLE TITLE")
points(train[train$ACTION==0,c(2,7)],col="black",pch=19)
abline(h=135000)
abline(h=255000)
```
#### Answer: 


### Similar plots for all variables!
```{r}
resource_acceptance <- function(column)
  {
  par(xpd=FALSE)
  plot(train[train$ACTION==1,c(2,column)],col="red",pch=19)
  points(train[train$ACTION==0,c(2,column)],col="black",pch=19)
  }

par(mfrow=c(3,3))
resource_acceptance(3)
resource_acceptance(4)
resource_acceptance(5)
resource_acceptance(6)
resource_acceptance(7)
resource_acceptance(8)
resource_acceptance(9)
resource_acceptance(10)
```


### Question: what is left if we eliminate the clear acceptance cases?
```{r}
rest <- subset(train,
              train$MGR_ID < 14000 &
              train$ROLE_CODE<128000 &
              train$RESOURCE < 150000 &
              (train$ROLE_TITLE < 150000 | train$ROLE_TITLE>290000) &
              train$ROLE_ROLLUP_2 < 135000 &
              train$ROLE_ROLLUP_2 > 50000 &
              train$ROLE_ROLLUP_1 < 122000 &
              train$ROLE_ROLLUP_1 > 100000 &
              (train$RESOURCE < 49000 | train$RESOURCE > 69000) &
              (train$ROLE_ROLLUP_2 > 128000 | train$ROLE_ROLLUP_2 < 124000) & 
              (train$MGR_ID < 9500 | train$MGR_ID > 13000) &
               train$ROLE_FAMILY_DESC > 50000
               )

resource_acceptance <- function(column)
{
  par(xpd=FALSE)
  plot(rest[rest$ACTION==1,c(2,column)],col="red",pch=19)
  points(rest[rest$ACTION==0,c(2,column)],col="black",pch=19)
}

par(mfrow=c(3,3))
resource_acceptance(3)
resource_acceptance(4)
resource_acceptance(5)
resource_acceptance(6)
resource_acceptance(7)
resource_acceptance(8)
resource_acceptance(9)
resource_acceptance(10)
```
#### Answer: We can eliminate almost half of points! But not sure what to do with the rest.

### Question: How does the distribution of features looks for TEST set?
```{r}
rest <- test[test$MGR_ID < 14000 &
            test$ROLE_CODE<128000 &
            test$RESOURCE < 150000 &   
            (test$ROLE_TITLE < 150000 | test$ROLE_TITLE>290000) &
            test$ROLE_ROLLUP_2 < 135000 &
            test$ROLE_ROLLUP_2 > 50000 &
            test$ROLE_ROLLUP_1 < 122000 &
            test$ROLE_ROLLUP_1 > 100000 &
            (test$RESOURCE < 49000 | test$RESOURCE > 69000) &
            (test$ROLE_ROLLUP_2 > 128000 | test$ROLE_ROLLUP_2 < 124000) & 
            (test$MGR_ID < 9500 | test$MGR_ID > 13000) &
            test$ROLE_FAMILY_DESC > 50000
             ,]

resource_acceptance <- function(column)
{
  par(xpd=FALSE)
  plot(rest[,c(2,column)],col="blue",pch=19)
}

par(mfrow=c(3,3))
resource_acceptance(3)
resource_acceptance(4)
resource_acceptance(5)
resource_acceptance(6)
resource_acceptance(7)
resource_acceptance(8)
resource_acceptance(9)
resource_acceptance(10)
```


## Ideas
* Make a simple logistic regression model to predict acceptance or not. 
* Try a classification tree!
* NO!!! Too many levels in factors! I need something simpler. How about i Choose first all factors with resource that only shows up once! 



